---
title: "Assignment7"
author: "Srishti Patil"
date: "2024-11-21"
output: html_document
---
1. Explain the output

2. Run Residual Analysis

3. Perform and Plot the forecast for the next five periods
4. Show the accuracy of your ARIMA model

```{r}
library(fpp)
library(fpp2)
library(forecast)
library(readxl)
#importing dataset
file_path <- "C:\\Users\\Srishti\\Desktop\\FALL24\\BF\\Assignment3\\Data_NewYork.xlsx"
dataFile <- read_excel(file_path)
head(dataFile)
timeSeriesData <- ts(dataFile$UnemployedLF, start=c(1976, 1), frequency=12)
plot(timeSeriesData)

#Perform and Plot the forecast for the next five periods
ndiffs(timeSeriesData)
tsdisplay(timeSeriesData)
timeseriesdiff1 <- diff(timeSeriesData, differences=1)
plot(timeseriesdiff1)
tsdisplay(timeseriesdiff1)
auto_fit <- auto.arima(timeSeriesData, trace=TRUE, stepwise = FALSE)
auto_fit
attributes(auto_fit)
plot(forecast(auto_fit,h=5,level=c(99.5)))
```


The best ARIMA model for the time series data was identifies as: ARIMA(0,1,2)(0,0,1)[12]. 
Explanations:
Model Components are ARIMA(0,1,2) which states:
p=0: No autoregressive terms.
d=1: Data is differenced once to achieve stationarity.
q=2: Two moving average terms.

Seasonal Component  are (0,0,1)[12] which states:
P=0: No seasonal autoregressive terms.
D=0: No seasonal differencing.
Q=1: One seasonal MA term.
[12]: Indicates seasonality occurs with a periodicity of 12.

Coefficients
The coefficients of the model are:
ma1=0.358
7: First-order moving average parameter.
ma2=−0.0919: Second-order moving average parameter.
sma1=−0.0655: Seasonal moving average parameter.

The output states that the model accounts for short-term trends using the two MA terms and one seasonal MA term and the seasonal patterns are identified based on the periodicity of 12.

```{r}

# Calculate the accuracy of the ARIMA model
model_accuracy <- accuracy(auto_fit)
print(model_accuracy)
```
We can see that RMSE and MAE are both low here. MAPE is very low which shows excellent relative accuracy. Hence, MAPE is the best in this case.

```{r}
#Run Residual Analysis
Acf(auto_fit$residuals)
Box.test(residuals(auto_fit), lag=20, type="Ljung")
plot.ts(residuals(auto_fit))
hist(auto_fit$residuals)
tsdiag(auto_fit)

```







-The residuals show no significant autocorrelation (from the ACF plot), indicating that the model effectively captures the dynamics of the data.
-The residuals are mostly random, with no visible trends or patterns in the time series plot.
-The histogram indicates a near-normal distribution of errors.
-The residuals' variance increases toward the later time periods, hinting at possible heteroscedasticity or the presence of outliers in specific periods.
